---
author: "Rik Ferreira Alves"
title: "Railways across the world"
subtitle: "Data from WorldBank"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

### Libraries

```{r message=FALSE}
library(tidyverse)
library(wbstats)
```

## Feature extraction

### Goods transported by railways (ton/km)

```{r}
railways_goods <- wb_data("IS.RRS.GOOD.MT.K6", mrnev = 1)

railways_goods <- railways_goods %>%
  arrange(-IS.RRS.GOOD.MT.K6) %>%
  rename(GOODS_T = IS.RRS.GOOD.MT.K6)
```

### People transported by railways (person/km)

```{r}
railways_people <- wb_data("IS.RRS.PASG.KM", mrnev = 1)

railways_people <- railways_people %>%
  arrange(-IS.RRS.PASG.KM) %>%
  rename(PEOPLE_T = IS.RRS.PASG.KM)
```

### Country population (thousands)

```{r}
population <- wb_data("SP.POP.TOTL", mrnev = 1)

population <- population %>%
  rename(POPULATION = SP.POP.TOTL)
```

### Country GDP (PPP)

```{r}
gdp_ppp <- wb_data("NY.GDP.PCAP.PP.CD", mrnev = 1)

gdp_ppp <- gdp_ppp %>%
  rename(GDP_PPP = NY.GDP.PCAP.PP.CD)
```

## Dataset construction

```{r}
dataset <- railways_goods %>%
  select(iso3c, GOODS_T) %>%
  full_join(
    railways_people %>%
      select(iso3c, PEOPLE_T),
    by = "iso3c"
  ) %>%
  full_join(
    population %>%
      select(iso3c, POPULATION),
    by = "iso3c"
  ) %>%
  full_join(
    gdp_ppp %>%
      select(iso3c, GDP_PPP),
    by = "iso3c"
  )
```

## Correlations

### People transported on goods transported

Although there are some exceptions, it is noticeable that there is some positive correlaetion between the goods transported and the people transported within countries.

```{r message=FALSE, warning=FALSE}
dataset %>%
  ggplot() +
  geom_point(aes(GOODS_T, PEOPLE_T)) +
  geom_smooth(aes(GOODS_T, PEOPLE_T), method = "lm", se = FALSE) +
  annotate("text", label = paste0("Pearson: ", format(cor(dataset$GOODS_T, dataset$PEOPLE_T, use = "complete"), digits = 2)), x = 10, y = 10000) +
  scale_x_log10() +
  scale_y_log10()
```

### Railway demand

#### Goods transported on GDP

```{r message=FALSE, warning=FALSE}
dataset %>%
  ggplot() +
  geom_point(aes(GDP_PPP, GOODS_T)) +
  geom_smooth(aes(GDP_PPP, GOODS_T), method = "lm", se = FALSE) +
  annotate("text", label = paste0("Pearson: ", format(cor(dataset$GDP_PPP, dataset$GOODS_T, use = "complete"), digits = 2)), x = 1e3, y = 1e5) +
  scale_x_log10() +
  scale_y_log10()
```

#### People transported on population

```{r message=FALSE, warning=FALSE}
dataset %>%
  ggplot() +
  geom_point(aes(POPULATION, PEOPLE_T)) +
  geom_smooth(aes(POPULATION, PEOPLE_T), method = "lm", se = FALSE) +
  annotate("text", label = paste0("Pearson: ", format(cor(dataset$POPULATION, dataset$PEOPLE_T, use = "complete"), digits = 2)), x = 1e3, y = 1e5) +
  scale_x_log10() +
  scale_y_log10()
```

